Data Cleaning

### 2024-11-26 study1/analysis/1_cleaning.qmd
Cleaning script updates

---
title: "1_cleaning"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(easystats)
library(ggplot2)
library(readr)
library(jsonlite)

df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

# Age demographics
p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=20) +
  # geom_line(aes(y = y), color = "orange", linewidth=2) +
  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age

# Education demographics
p_edu <- df |>
  mutate(Education = fct_relevel(Education, "Other", "Elementary school", "High school", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

# Discipline demographics
p_dis <- df |>
  ggplot(aes(x = Discipline)) +
  geom_bar(aes(fill = Discipline)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Discipline", y = "Number of Participants", subtitle = "Participants per discipline") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

#
```

>>>>

---
title: "Data Cleaning"
format: 
  html:
      code-fold: true
editor: visual
---

```{r}
#/ output

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)

df <- read_csv("C:/Users/maisi/OneDrive/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

# df <- read_csv("C:/Users/emmal/OneDrive - University of Sussex/GitHub/Placement/InteroceptionScale/study1/data/rawdata_participants.csv")

```

## Age and Gender

```{r}
#/ code-fold: TRUE

p_age <- estimate_density(df$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=df, aes(x = Age, fill=Gender), bins=28) +
  # geom_line(aes(y = y), color = "orange", linewidth=2) +
  geom_vline(xintercept = mean(df$Age), color = "#ca0020", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B", "Female"= "#8EB69B", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age and Gender", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age and gender") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

## Education

```{r}
#/ code-fold: TRUE

# Education 
# did not add education disiplines

p_edu <- df |>
  mutate(Education = fct_relevel(Education, "Secondary (technical) School", "Elementary School", "High School", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  # scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
   scale_fill_manual(values = c("Secondary (technical) school"="#0571B0","High school"= "#d01c8b", "Bachelor"="#FDB863", "Master"="#E66101", "Doctorate" =  "#ca0020")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5)),
    axis.title.x = element_blank()
  )

p_edu

# Student

table(df$Student)
```

## Ethinicity

```{r}
#/ code-fold: TRUE
# Ethnicity

p_eth <- df |>
  mutate(Ethnicity = fct_relevel(Ethnicity,"East Asian", "Mixed", "White", "Black", "persian", "South Asian", "Hispanic/Latino")) |>
  ggplot(aes(x = Ethnicity)) +
  geom_bar(aes(fill = Ethnicity)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Ethnicity", y = "Number of Participants", subtitle = "Participants per Ethnicity") +
  theme_modern(axis.title.space = 15) +
  scale_fill_manual(values = c("East Asian"="#0571B0","Mixed"= "#d01c8b", "White"="#FDB863", "Black"="#E66101", "persian" =  "#ca0020", "South Asian"="#5E3C99", "Hispanic/Latino"="#92C5DE")) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.6)),
    axis.title.x = element_blank())

p_eth

```

## Residence

```{r}

# Country participants are currently living

p_cou <- df |>
  mutate(Country = fct_relevel(Country,"United Kingdom", "Ireland", "USA", "Australia")) |>
  ggplot(aes(x = Country)) +
  geom_bar(aes(fill = Country)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Country of Residence", y = "Number of Participants", subtitle = "Participants per Country they are currently living") +
  theme_modern(axis.title.space = 15) +
  scale_fill_manual(values = c("United Kingdom"="#ca0020","Ireland"= "#d01c8b", "United States"="#FDB863", "Australia"="#E66101"))+
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.7)),
    axis.title.x = element_blank())

p_cou

```

```{r}

# <!--Do it-->

p_country <- dplyr::select(df, region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n)) +
  scale_fill_gradientn(colors = c("#E66101", "#ca0020", "#d01c8b")) +
  labs(fill = "N") +
  theme_void() +
  labs(title = "Country of Residence", subtitle = "Number of participants by country")  +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2))
  )
p_country
```

```{r}

table(df$Country)
```

```{r}

dplyr::select(df, region = Country) |>
  group_by(region) |>
  summarize(n = n()) |>
  right_join(map_data("world"), by = "region") |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = n))

```

```{r}

p_feed <- estimate_density(df$Experiment_Enjoyment) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_bar(data=df, aes(x = Experiment_Enjoyment, fill = Condition), bins=28) +
  scale_fill_manual(values = c("Dimensions"= "#E66101","Domains" = "#ca0020", "Random"="#d01c8b")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Experiment Enjoyment and Condition", y = "Number of Participants", color = NULL, x = "Ratings of enjoyment on a scale 1-5", subtitle = "Distribution of Participants' Experiment Enjoyment and Condition") +
  # geom_label(data = data.frame(x = mean(df$Experiment_Enjoyment) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Experiment_Enjoyment)))) +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_feed

  #geom_line(aes(y = y), color = "orange", linewidth=2) +
#geom_vline(xintercept = mean(df$Experiment_Enjoyment), color = "#ca0020", linewidth=1.5) +
```
